### 2017-09-02 paper/paper/paper.tex
Rudolf's edits. No Kohler, proportionality, several wording upgrades

\documentclass[10pt]{article}
\usepackage[OE]{express}
\usepackage{float}
\usepackage{booktabs}
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
\providecommand{\mb}[1]{\mathbf{#1}}
\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
\providecommand{\bs}[1]{\boldsymbol{#1}} 
\providecommand{\intinf}{\int_{-\infty}^{\infty}}

\begin{document}
\title{Single-fluorophore orientation determination with multiview polarized
  illumination: modeling and microscope design}

\author{Talon Chandler,\authormark{1,*} Shalin Mehta,\authormark{1,2,3} Hari
  Shroff,\authormark{4,5} Rudolf Oldenbourg,\authormark{2, 6} and Patrick J. La
  Rivi\`ere\authormark{1, 5}}
\address{\authormark{1}University of Chicago, Department of Radiology, Chicago, Illinois 60637, USA\\
  \authormark{2}Marine Biological Laboratory, Bell Center, Woods Hole, Massachusetts 02543, USA\\
  \authormark{3}(present address) Chan Zuckerberg Biohub, San Francisco, California 94158, USA\\
  \authormark{4}Section on High Resolution Optical Imaging, National Institute
  of Biomedical Imaging and Bioengineering, National Institutes of Health,
  Bethesda, Maryland 20892, USA\\
  \authormark{5}Marine Biological Laboratory, Whitman Center, Woods Hole,
  Massachusetts 02543, USA\\
  \authormark{6}Brown University, Department of Physics, Providence, Rhode
  Island 02912, USA}
\email{*talonchandler@talonchandler.com} %% email address is required

%%%%%%%%%%%%%%%%%%% abstract and OCIS codes %%%%%%%%%%%%%%%%
\begin{abstract*}
  We investigate the use of polarized illumination in multiview microscopes for
  determining the orientation of single fluorescent molecules. First, we relate
  the orientation of single molecules to measurable intensities in multiview
  microscopes and develop an information-theoretic metric---the solid-angle
  uncertainty---to compare the ability of multiview microscopes to estimate the
  orientation of single molecules. Next, we compare a broad class of microscopes
  using this metric---single- and dual-view microscopes with varying
  illumination polarization, illumination numerical aperture (NA), detection NA,
  obliquity, asymmetry, and exposure. We find that multiview microscopes are
  better at determining the orientation of single molecules than single-view
  microscopes. We also find that choosing a small illumination NA and a large
  detection NA are good design choices, that multiview microscopes can benefit
  from oblique illumination and detection, and that asymmetric NA microscopes
  can benefit from exposure asymmetry.
\end{abstract*}

\ocis{(110.0110) Imaging systems; (180.2520) Fluorescence microscopy; (180.0180)
  Microscopy; (180.6900) Three-dimensional microscopy; (130.5440)
  Polarization-selective devices; (260.5430) Polarization.}

%%%%%%%%%%%%%%%%%%%%%%% References %%%%%%%%%%%%%%%%%%%%%%%%%
%% Comment before submission
\bibliography{paper}
\bibliographystyle{osajnl}

%% Copy .bbl and uncomment these before submission
% \begin{thebibliography}{99}
% \bibitem{gallo99} K. Gallo and G. Assanto, ``All-optical diode based on second-harmonic generation in an asymmetric waveguide,'' \josab {\bfseries 16}(2), 267--269 (1999).
% \end{thebibliography}

\section{Introduction}
The orientation of single molecules is a valuable reporter for
biological processes. By imaging biological samples with fluorescent probes that
are fixed to known structures, researchers have studied the dynamics of motor
proteins \cite{peterman2001, forkey2003}, DNA \cite{backer2016}, actin
\cite{mehta2016}, septin \cite{demay2011, mcquilken2017}, and membranes
\cite{anantharam2010}. The number of techniques available to biologists for
measuring the orientation of single molecules in live cells is continually
expanding. A recent review \cite{backlund2014} identified three major categories
of single-molecule orientation determination techniques:

\textbf{Spatial/angular variation of emission} techniques image single molecules
and use the distribution of intensity in the back focal plane \cite{lieb2004} or
in the image plane \cite{backer2014} to estimate their
orientation. Information-theoretically optimal techniques in this category use
defocusing or phase masks in the back focal plane to encode orientation and
position information in the image \cite{agrawal2012}. These techniques can
estimate the orientation of molecules in all orientations and only require a
single frame of intensity measurements, but they require complex reconstruction
algorithms, expensive high numerical aperture (NA) optics, and sensitive
detectors. Furthermore, these techniques spread the signal from a single
molecule over a large area on the detector, so they suffer from a poor
signal-to-noise ratio \cite{toprak2006}.

\textbf{Spatial variation of illumination polarization} techniques vary the
polarization and intensity in the illumination path to excite molecules in
specific orientations in the focal volume \cite{debarre2004}. Although these
techniques can accurately estimate the orientation of molecules in all
orientations, they require scanning which makes acquisition time a concern in
live-cell imaging.

\textbf{Polarized illumination or detection} techniques vary the polarization of
light in the illumination or detection paths and exploit the anisotropic
absorption and emission patterns of single molecules to estimate their
orientation. By illuminating and detecting from the entire sample, not just the
focal volume, these techniques can estimate the orientation of many molecules in
parallel without scanning. These techniques are easy to implement---changing the
illumination or detection polarization is simple with a polarization splitter
\cite{mehta2016} or universal compensator/polarizer \cite{shribak2003}---and the
reconstruction methods are straightforward \cite{fourkas2001, mehta2016,
  backer2016}. Also, these techniques keep the signal from a single molecule
localized on the detector, so they have a signal-to-noise ratio advantage over
defocusing and phase masking techniques. The main drawback is that these
techniques require several intensity measurements under different polarization
orientations to estimate the orientation of a single molecule. Despite the
requirement for multiple measurements, polarized illumination or detection
methods are good choices for live cell imaging because of their combination of
speed and ease of implementation.

To develop a polarized illumination or detection technique we need to choose a
method for varying the polarization and place it in the illumination or
detection path. The simplest and most inexpensive way to vary the polarization
is with a division-of-time (DoT) polarizer using a rotating polarizer or a
universal compensator/polarizer. Other polarization methods
(divison-of-amplitude, division-of-focal-plane) are more expensive and require
additional alignment or interpolation \cite{tyo2006}. The simplicity of DoT
polarization comes at the expense of temporal resolution---intensity
measurements are made one after the other which increases the acquisition
time. If we place the DoT polarizer in the detection path, we will block
valuable photons from reaching the detector which leads to unnecessary exposure
to the sample \cite{demay2011}. By placing the DoT polarizer in the illumination
arm we minimize exposure to the sample. We focus on DoT polarized illumination
in this paper because of its combination of simplicity and low sample exposure.

All single molecule orientation determination methods suffer from some degree of
anisotropic orientation uncertainty, i.e., some molecular orientations cannot be
determined as precisely as others. In some cases the orientation cannot be
determined at all---the forward model can be degenerate for specific dipole
orientations \cite{fourkas2001, lu2008}. We feel that the importance of
isotropic orientation uncertainty has been underappreciated. To our knowledge,
the only authors who have considered anisotropic orientation uncertainty have
only analyzed a subset of molecular orientations \cite{agrawal2012}. Our view is
that an ideal technique for determining the orientation of single molecules can
reconstruct the orientation with a small and nearly uniform uncertainty for all
molecule orientations.

Recently, there has been increasing interest in multiview microscopy techniques
for biological imaging \cite{wu2013, keller2015, wu2016, wu2017}. Multiview
microscopes offer two major advantages over single-view microscopes. First,
multiview microscopes can achieve nearly isotropic resolution compared to the
anisotropic resolution of single-view microscopes whose axial resolution is
reduced by at least a factor of two compared to lateral resolution. Second,
multiview microscopes can use light-sheet illumination to reduce phototoxicity
while reusing parts of the optical path for both illumination and detection. For
example, if two orthogonal objectives are focused on the same point, then the
objectives can alternate roles as the light-sheet illumination path and as the
detection path. Light-sheet illumination offers a major reduction of
phototoxicity because a light sheet only illuminates in-focus regions while
other widefield and confocal approaches inevitably illuminate out-of-focus
regions. Together, these advantages make multiview microscopes good candidates
for imaging live biological specimens for long periods.

In this work we explore the use of polarized illumination in multiview
microscopes for determining the orientation of single molecules. Existing
multiview microscopes can easily be outfitted with fast-switching polarizers
that do not degrade image quality, so polarized illumination is a natural way to
augment multiview microscopes for measuring the orientation of single
molecules. Furthermore, multiview microscopes can achieve nearly isotropic
resolution while delivering selective illumination. We find that multiview
microscopes also provide a small and uniform orientation uncertainty for single
molecules in all orientations. We also find that single-view microscopes have
symmetries that can be broken by adding an second view. This reduction of
degeneracy allows multiview microscopes to uniquely determine the orientation of
single molecules.

In section \ref{methods} we develop the theory required to model polarized
illumination microscopes, and we develop metrics to compare microscope
designs. In section \ref{results} we compare the results for a wide range of
multiview microscope designs. Finally, in section \ref{discussion} we discuss
the results and their impact on polarized multiview microscope design.

\section{Methods}\label{methods}
In this section we develop a method to compare multiview polarized illumination
microscopes for the task of reconstructing molecular orientations. In sections
\ref{excitation}--\ref{forward} we develop a model for the intensity collected
with a fixed illumination and detection configuration as a function of the
molecule orientation. In section \ref{designs} we list a variety of ways to
combine these intensity measurements into experimentally realizable polarized
illumination microscopes. Finally, in section \ref{metrics} we develop metrics
that we will use to compare these microscopes.

We use roman type for scalars, e.g., $\phi$, $\theta$; bold lowercase type for
vectors, e.g., $\mb{r}, \bs{\mu}$; hats for unit vectors, e.g.,
$\mh{r}, \hat{\bs{\mu}}$; and bold capital type for matrices, e.g.,
$\mb{R}, \mb{F}$.

\subsection{Absorption Efficiency}\label{excitation}
In this section we will calculate the \emph{absorption efficiency} of a single
molecule---the fraction of the incident power that is absorbed by the
molecule. Our approach is inspired by Fourkas \cite{fourkas2001}, but here we
calculate the absorption efficiency instead of the detection efficiency.

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{frames}
\caption{Coordinate systems for a) the absorption dipole moment
  $\hat{\bs{\mu}}_{\text{abs}}$, b) the polarizer transmission axis
  $\hat{\mb{e}}_{\text{exc}}$, and c) the dummy integration vector
  $\hat{\mb{r}}$. d) The scene consists of a single molecule at the focal point
  of an objective lens with a polarizer in the aperture plane.}
\label{fig:coordinates}
\end{figure}

We consider a single molecule with a fixed absorption dipole moment
$\hat{\bs{\mu}}_{\text{abs}}$ that we express in spherical coordinates $\Theta$
and $\Phi$ (see Figure \ref{fig:coordinates}a)) as
\begin{align}
  \hat{\bs{\mu}}_{\text{abs}}(\Theta, \Phi) = \sin\Theta\cos\Phi\hat{\mb{x}} + \sin\Theta\sin\Phi\hat{\mb{y}} + \cos\Theta\hat{\mb{z}} \label{eq:mu}.
\end{align}
We place the molecule in the focal plane of an ideal, aplanatic,
polarization-preserving objective lens with its optical axis aligned with the
$\mh{z}$ axis. Next, we place a polarizer in the aperture plane with a variable
transmission axis $\hat{\mb{p}}_{\text{exc}}$ that we express as
\begin{align}
 \hat{\mb{p}}_{\text{exc}}(\phi_{\text{exc}}) = \cos\phi_{\text{exc}}\hat{\mb{x}} +
\sin\phi_{\text{exc}}\hat{\mb{y}},\label{eq:excfield}
\end{align}
where $\phi_{\text{exc}}$ is the angle between the transmission axis of the
polarizer and the positive $\mh{x}$ axis (see Figure
\ref{fig:coordinates}b)). Finally, we place a spatially incoherent and spatially
uniform light source (or its image) in the aperture plane to illuminate the
focal plane. In this geometry each point in the sample is illuminated by a range
of angles that can be changed by adjusting the radius of the stop in the
aperture plane. We denote the set of illuminated angles by $\Omega$ (see Figure
\ref{fig:coordinates}d)). If we use unit vectors $\hat{\mb{r}}$ expressed in
spherical coordinates $\theta$ and $\phi$ (see Figure \ref{fig:coordinates}c))
as
  \begin{align}
   \hat{\mb{r}}(\theta, \phi) = \sin\theta\cos\phi\hat{\mb{x}} + \sin\theta\sin\phi\hat{\mb{y}} + \cos\theta\hat{\mb{z}}, 
  \end{align}
  then the set of angles $\Omega$ can be expressed as
  \begin{align}
  \Omega = \{\phi, \theta\ |\ 0 < \phi \leq 2\pi, 0 < \theta \leq \alpha\},
  \end{align}
  where $\alpha$ is the angle between the optical axis and the most oblique
  illuminating ray. Equivalently, $\alpha$ can be expressed in terms of NA using
  \begin{align}
    \text{NA} = n\sin\alpha, 
  \end{align}
  where $n$ is the index of refraction of the sample medium. Note that the NA of
  the objective is always greater than or equal to the illumination NA because
  we can underfill the aperture plane.

  The objective focuses the light by applying a position-dependent rotation to
  the electric field which we model by multiplying the incident electric field
  $\hat{\mb{p}}_{\text{exc}}$ by a position-dependent rotation matrix
  \begin{align}
  \mb{R}(\mh{r}) &= \begin{bmatrix} \cos\theta\cos^2\phi + \sin^2\phi & (\cos\theta -1)\sin\phi\cos\phi & -\sin\theta\cos\phi\\ (\cos\theta - 1)\sin\phi\cos\phi & \cos\theta\sin^2\phi + \cos^2\phi & -\sin\theta\sin\phi \\ \sin\theta\cos\phi& \sin\theta\sin\phi & \cos\theta \end{bmatrix} \label{eq:rot_mat}.
  \end{align}

  To find the absorption efficiency $\eta_{\text{abs}}$ of a single molecule if
  it were illuminated by a single ray, we take the unit incident electric field
  $\hat{\mb{p}}_{\text{exc}}$, take the dot product with the absorption dipole
  moment, then take the modulus squared
    \begin{align}
      \eta_{\text{abs}}^{\text{ray}} &= |\hat{\bs{\mu}}_{\text{abs}}\cdot\mh{p}_{\text{exc}}|^2. \label{eq:singleray}
    \end{align}

    To find the absorption efficiency of a single molecule, we integrate over
    all illumination rays and divide by the total incident power which gives the
    vector expression
\begin{align}
  \eta_{\text{abs}} = \frac{\int_{\Omega}d\mh{r}|\hat{\bs{\mu}}_{\text{abs}}\cdot\mb{R}(\mh{r})\mh{p}_{\text{exc}}|^2}{\int_{\Omega}d\mh{r}}\label{eq:abs}. 
\end{align}

We substitute equations \ref{eq:mu}--\ref{eq:rot_mat} into equation
\ref{eq:abs}, evaluate the integrals, and simplify to express the absorption
efficiency in scalar notation as
\begin{align}
  \eta_{\text{abs}} &= D\{A + B\sin^{2}{\Theta} + C\sin^{2}{\Theta} \cos{[2 (\Phi - \phi_{\text{exc}}})]\}\label{eq:scalarabs}
\end{align}
where
\begin{subequations}
\begin{align}
  A &= \frac{1}{4} - \frac{3}{8} \cos{\alpha } + \frac{1}{8} \cos^{3}{\alpha }\\
  B &= \frac{3}{16} \cos{\alpha } - \frac{3}{16} \cos^{3}{\alpha }\\
  C &= \frac{7}{32} - \frac{3}{32} \cos{\alpha } - \frac{3}{32} \cos^{2}{\alpha } - \frac{1}{32} \cos^{3}{\alpha}\\
  D &= \frac{4}{3(1 - \cos\alpha)}.
\end{align}\label{eq:coefficients}%
\end{subequations}

We can extend these results to light-sheet illumination created by scanning a
laser. If we shine a narrow, collimated, laser beam onto a galvanometer in the
aperture plane and scan the galvanometer, we create a weakly focused light sheet
in the sample. If we scan the beam in the sample with a velocity much less than
the beam waist radius divided by the coherence time, then the coherence of the
laser is rendered ineffectual by the scanning---the light sheet is effectively
composed of many incoherent beams. If we also ensure that the laser beam is
weakly focused, we can ignore the excitation caused by the small fields aligned
with the optical axis. Under slowly-scanned and weakly-focused light-sheet
illumination, molecules in the sample will be excited as if a single plane wave
was incident. Therefore, we can find the absorption efficiency of a molecule in
this case by taking the limit of equation \ref{eq:scalarabs} as
$\alpha \rightarrow 0$ (or by plugging equation \ref{eq:mu} and
\ref{eq:excfield} into equation \ref{eq:singleray}) giving
\begin{align}
  \eta_{\text{abs}}^{\text{ray}} &= \sin^2\Theta\cos^2(\Phi - \phi_{\text{exc}}),
\end{align}
which is recognizable as Malus' law generalized to three dimensions. This means
that light-sheet illumination is approximately equivalent to low-NA illumination
with respect to the molecular orientations they excite. Notice that the
absorption efficiency can take its maximum range of values
($\eta_{\text{abs}} \in [0, 1]$) when $\alpha=0$.

\subsection{Detection Efficiency}\label{detection}
Fourkas calculated the detection efficiency of a single molecule when an ideal
objective with its optical axis aligned with the $\mh{z}$ axis is focused on the
molecule and a polarizer is placed in the aperture plane\cite{fourkas2001}. We
use his expressions to calculate the detection efficiency without a polarizer as
\begin{align}
  \eta_{\text{det}} = 2(A + B\sin^2\Theta). \label{eq:scalardet}
\end{align}
The detection efficiency only depends on $\Theta$, not $\Phi$, because there is
no detection polarizer. We have assumed that the absorption dipole moment is
equal to the emission dipole moment.

Note that $A$ and $B$ in equation \ref{eq:coefficients} are a factor of
$\frac{3}{2}$ larger than the expressions given by Fourkas. We found that
Fourkas' expressions were incorrectly normalized (the limit of
$\eta_{\text{det}}$ as $\alpha\rightarrow \pi/2$ should be $\frac{1}{2}$), and
the extra factor of $\frac{3}{2}$ corrects the error. This correction does not
affect Fourkas' orientation reconstruction, but it does mean that Fourkas'
algorithm underpredicts the total emitted intensity by a factor of
$\frac{3}{2}$.

Also note that the detection efficiency can take its maximum range of values
when $B$ takes its maximum value, i.e. when
$\alpha=\arccos\left(\frac{1}{\sqrt{3}}\right) \approx 54.7^{\circ}$. When
$\alpha = 90$ the detection efficiency is $\frac{1}{2}$ for all fluorophore
orientations and when $\alpha = 0$ the detection efficiency is zero for all
fluorophore orientations, so we expect there to be an intermediate value of
$\alpha$ where the detection efficiency takes its maximum range.

\subsection{Oblique Illumination and Detection}\label{oblique}
In sections \ref{excitation} and \ref{detection} we assumed that both the
illumination and detection objectives had $\mh{z}$ aligned optical axes, i.e.,
we used epi- or transmitted light illumination. To extend the forward model to
oblique optical axes which allows for non-coincident illumination and detection
objectives, we express the dipole orientation in a rotated coordinate frame
using the following expressions
\begin{align}
    \Theta' &= \arccos\left(\sin\psi\cos\Phi\sin\Theta + \cos\psi\cos\Theta\right)\label{eq:thetap}\\
  \Phi' &=
          \begin{cases}
            \arccos\left(\dfrac{\cos\psi\cos\Phi\sin\Theta - \sin\psi\cos\Theta}{\sqrt{1 - (\sin\psi\cos\Phi\sin\Theta + \cos\psi\cos\Theta)^2}}\right) \ \ \ &0 \leq \Phi < \pi  \\
            -\arccos\left(\dfrac{\cos\psi\cos\Phi\sin\Theta - \sin\psi\cos\Theta}{\sqrt{1 - (\sin\psi\cos\Phi\sin\Theta + \cos\psi\cos\Theta)^2}}\right) \ \ \ &-\pi \leq \Phi < 0\\
          \end{cases}\label{eq:phip}
\end{align}
where $\psi$ is the angle of a right handed rotation about the $\mh{y}$ axis
that maps the $\mh{z}$ axis onto the new optical axis.

\subsection{Intensity Measurements}\label{forward}
The detected intensity is proportional to the product of the sample exposure
$I_{\text{exp}}$, the absorption efficiency, and the detection efficiency
\begin{align}
  I \propto I_{\text{exp}}\eta_{\text{abs}}\eta_{\text{det}}\label{eq:forward_frame},
\end{align}
where the proportionality depends on the exposure time, the molecular quantum
efficiencies, and other constant factors. Equation \ref{eq:forward_frame} is the
forward model for a single intensity measurement. Figure
\ref{fig:single-frame} shows the efficiencies in three representative
geometries. Note that in all cases there are multiple orientations that give the
same intensity measurement. This means that a single intensity measurement does
not give us enough information to reconstruct the orientation of a single
molecule. Next, we will consider combining several intensity measurements to
create complete polarized illumination microscopes.

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{single-frame}
\caption{Representative examples of single intensity measurements. Black dots
  indicate where the Cartesian unit vectors intersect the unit sphere. \newline \newline \textbf{Columns left to right:} 1) schematics where
  the solid line encloses the illumination solid angle, the dashed line encloses
  the detection solid angle, and the arrow indicates the transmission axis of
  the illumination polarizer; 2) the absorption efficiency, see equation
  \ref{eq:scalarabs}; 3) the detection efficiency, see equation
  \ref{eq:scalardet}; (4) the total efficiency, the product of the absorption
  and detection efficiencies.\newline \newline \textbf{Rows top to bottom:} 1)
  coincident illumination ($\text{NA} = 1.1$ with $x$-polarized light) and
  detection (NA = 1.1); 2) non-coincident orthogonal illumination
  ($\text{NA} = 0.8$) and detection ($\text{NA} = 0.8$); 3) non-coincident
  135${}^{\circ}$-separated illumination ($\text{NA} = 0.8$) and detection
  ($\text{NA} = 0.8$). All simulations use $n=1.33$.}
  \label{fig:single-frame}
\end{figure}

\subsection{Polarized Illumination Microscope Configurations}\label{designs}
One way to collect multiple intensity measurements is to add a universal
compensator/polarizer to the illumination arm and rapidly select the incident
polarization by changing $\phi_{\text{exc}}$ \cite{shribak2003}. All of the
microscopes we will consider in this paper use four polarizations per
illumination path with the polarization orientations separated by
45${}^{\circ}$.

We also consider several dual-view designs. We evaluate dual-view designs that
allow for illumination and detection from both objectives (see \cite{wu2013} for
a dual-view design). In all cases we consider the effect of varying the
illumination and detection numerical aperture. We also consider asymmetric NA
designs (see \cite{wu2017} for an asymmetric NA design) and the effect of
asymmetric sample exposures. Note that we use the word \emph{symmetry} to refer
to dual-view designs with identical objectives and sample exposures and the word
\emph{asymmetry} to refer to dual-view designs with objectives that have a
different NA or sample exposure.

All multiview microscopes are subject to steric constraints---the objectives
must not collide with the cover slip or each other. We only consider microscope
designs that meet these criteria.

\subsection{Evaluation Metrics}\label{metrics}
Our goal is to evaluate the ability of a microscope design to estimate the
orientation of a single molecule (via parameters $\Theta$ and $\Phi$) from
intensity data. A common way to evaluate the ability to estimate the parameters
$\Theta$ and $\Phi$ from the data is to calculate the Cram\'er-Rao lower bound
(CRLB) for each parameter \cite{kay1993}. The CRLBs are given by the diagonal
elements of the inverse of the Fisher information matrix, and they give the
minimum variance of an unbiased estimator for each parameter. For each
microscope design and molecule orientation we calculate the Fisher information
matrix. We assume that the detected intensities are Poisson distributed, so the
Fisher information matrix is given by
\begin{align}
  \mb{F} &= \sum_{k=1}^N \frac{1}{I_k}
  \begin{bmatrix}
    \dfrac{\partial I_k}{\partial \Theta}\dfrac{\partial I_k}{\partial \Theta}&\dfrac{\partial I_k}{\partial \Theta}\dfrac{\partial I_k}{\partial \Phi}\\\\
    \dfrac{\partial I_k}{\partial \Theta}\dfrac{\partial I_k}{\partial \Phi}&\dfrac{\partial I_k}{\partial \Phi}\dfrac{\partial I_k}{\partial \Phi}\\    
  \end{bmatrix}
\end{align}
where $I_k$ is the $k$th intensity measurement in a microscope that uses $N$
intensity measurements. Agrawal et. al. used the square root of the product of
the CRLBs multiplied by the Jacobian determinant,
$\sin\Theta\sqrt{\mb{F}^{-1}_{0,0}\mb{F}^{-1}_{1,1}}$, to find the area of
uncertainty in parameter space \cite{agrawal2012}.

CRLBs and the associated area of uncertainty are parametrization dependent,
i.e., if we choose a different coordinate system these metrics will change. We
would like to compare microscope designs without choosing a parametrization, so
instead we use
\begin{align}
  \sigma_{\Omega} &\equiv \sin\Theta\sqrt{\text{det}\{\mb{F}^{-1}\}}
\end{align}
as our evaluation metric. We calculate the determinant of the inverse Fisher
information matrix---a parametrization-independent value called the
\emph{generalized variance} \cite{anderson1958}---take the square root, then
multiply by the Jacobian determinant, $\sin\Theta$. We call $\sigma_{\Omega}$
the \emph{solid-angle uncertainty} because it has units of steradians and is a
parametrization independent measure of the orientation uncertainty.

For each microscope we calculate $\sigma_{\Omega}$ at 10,000 approximately
equally spaced points on the unit sphere. A desirable microscope design will
have a small solid-angle uncertainty that is uniform for all molecule
orientations. The most straightforward way to find the location and scale of the
solid-angle uncertainty is to use the mean and variance, respectively. However,
the solid-angle uncertainty can take extremely large values, e.g., when a
fluorescent dipole is aligned with the optical axis of a single excitation
objective. Such large values can change the mean and variance dramatically
depending on the sample points, so we use the median and median absolute
deviation (MAD)---the median of the data's absolute difference from the
median---as robust alternatives to the mean and variance.

To compare microscopes with different numbers of intensity measurements fairly,
we kept the total exposure to the sample constant by choosing
$I_{\text{exp}} = \frac{4000}{N}$. Note that $I_{\text{exp}}$ is a measure of
exposure to the sample, not the detected intensity or the total intensity
emitted by the molecule.

We also use the Fisher information matrix to calculate uncertainty ellipses for
each microscope and fluorophore orientation. The general equation of an ellipse
with axis radii $a$ and $b$ and rotation angle $\gamma$ is
\begin{align}
  \frac{(x\cos\gamma + y\sin\gamma)^2}{a^2} + \frac{(x\cos\gamma - y\sin\gamma)^2}{b^2} =1. 
\end{align}
We calculate the ellipse parameters in terms of the Fisher information matrix
using the following set of equations
\begin{align}
  \sigma_x &= \sqrt{\mb{F}^{-1}_{0,0}} \label{eq:first}\\
  \sigma_y &= \sin\Theta\sqrt{\mb{F}^{-1}_{1,1}} \label{eq:second}\\ 
  \sigma_{xy} &= \sin\Theta\sqrt{\mb{F}^{-1}_{1,0}} \label{eq:third}\\  
  a^2 &= \frac{\sigma_x^2 + \sigma_y^2}{2} + \sqrt{\frac{(\sigma_x^2 - \sigma_y^2)^2}{4} + \sigma_{xy}^2}\\
  b^2 &= \frac{\sigma_x^2 + \sigma_y^2}{2} - \sqrt{\frac{(\sigma_x^2 - \sigma_y^2)^2}{4} + \sigma_{xy}^2}\\
  \gamma &= \frac{1}{2}\text{arctan}\left(\frac{2\sigma_{xy}}{\sigma_x^2 - \sigma_y^2}\right). \label{eq:last}
\end{align}
Equations \ref{eq:first}--\ref{eq:last} are identical to those found in
\cite{coe2009} with an additional factor of $\sin\Theta$ on equations
\ref{eq:second} and \ref{eq:third} to account for the non-Euclidean geometry of
the sphere. Instead of plotting ellipses with widely varying radii on the same
figure, we chose a constant major axis radius. This means that the uncertainty
ellipses indicate the relative size of uncertainty in different directions, not
the absolute size of the uncertainty.

The solid-angle uncertainty and the uncertainty ellipse are good local measures
of orientation uncertainty, but they are insensitive to degeneracies with
distant dipole orientations. For example, all of the intensity measurements in
this paper are identical under inversion of the dipole orientation
($I(\Theta, \Phi) = I(\Theta + \pi, \Phi)$ in equations \ref{eq:scalarabs},
\ref{eq:scalardet}, and \ref{eq:forward_frame}), but the solid-angle uncertainty
and the uncertainty ellipse give no indication of this degeneracy. For this
reason we also use the number of degeneracies as a measure of a micrsocope's
ability to estimate the orientation of a dipole from intensity measurements. A
small number of degeneracies is important for a microscope to find the
orientation of a molecule---a twofold degenerate microscope can find the
orientation of a molecule down to the hemisphere while a fourfold symmetric
microscope can only find the orientation down to the quadrant.

To calculate and display these evaluation metrics we use NumPy \cite{walt2011}
for numerical computation, SymPy \cite{meurer2017} for symbolic computation,
Matplotlib \cite{hunter2007} for high-level plotting, VisPy \cite{vispy} for 3D
rendered plots, and Asymptote \cite{bowman2008} for 3D line plots.

\section{Results}\label{results}
Figure \ref{fig:single-arm} shows our results for single-view designs where a
single objective is used for illumination and detection. We swept through the
illumination and detection NA while keeping the sample exposure constant, and we
found that the lowest median and MAD of the solid-angle uncertainty occurs with
a small illumination NA and a large detection NA where
$\alpha \approx 65^{\circ}$ (see Figure \ref{fig:single-arm}e) at
NA${}_{\text{det}} \approx 1.2$). A small illumination NA maximizes the range of
absorption efficiencies, while a large detection NA optimizes the number of
detected photons and the range of the detection efficiencies. The optimal
detection NA is slightly larger than $\alpha = 57.4^{\circ}$ because increasing
the detection NA reduces the effect of shot noise. Note the relative importance
of illumination and detection NAs---reducing the illumination NA (an inexpensive
modification requiring underfilling the aperture plane) improves the solid-angle
uncertainty much less than increasing the detection NA (an expensive
modification requiring a higher-NA objective). Also note in Figure
\ref{fig:single-arm}c) that single-view designs suffer from high orientation
uncertainty when the molecules are oriented along the optical axis (the
molecules are not efficiently excited or detected), near the transverse plane
(it is ambiguous whether the molecule is oriented above or below the transverse
plane, see Figure \ref{fig:single-arm}b)), and near the polarizer orientations
(it is ambiguous whether the molecule is on either side of the polarizer
orientation, see Figure \ref{fig:single-arm}b)).

The single-view microscope in Figure \ref{fig:single-arm} is fourfold degenerate
($I(\Theta, \Phi) = I(\Theta + \pi, \Phi) = I(\Theta, \Phi + \pi) = I(\Theta +
\pi, \Phi+\pi)$ in equations \ref{eq:scalarabs}, \ref{eq:scalardet}, and
\ref{eq:forward_frame}). In addition to the usual inversion degeneracy,
single-view microscopes suffer from a degeneracy when a dipole is reflected
through the transverse plane. Evidence of this degeneracy is found in Figures
\ref{fig:single-arm}b) and \ref{fig:single-arm}c) which show the ambiguity of
whether the molecule is oriented above or below the transverse plane. Near the
transverse plane this ambiguity is evident in the high solid-angle uncertainty,
but far from the transverse plane evidence of the ambiguity is lost. Adding a
second view and taking eight intensity measurements instead of four removes this
degeneracy (see equations \ref{eq:thetap} and \ref{eq:phip}), so the dual-view
microscopes we consider in this paper are only twofold degenerate. This means
that while dual-view microscopes can find the orientation of a molecule up to
the hemisphere, while single-view microscopes can only find the orientation of a
molecule up to the quadrant.

\begin{figure}[htbp]
\centering\includegraphics[width=\textwidth]{single-arm}
\caption{Single-view microscope with varying illumination and detection NA. a)
  Schematic of a single-view four-polarization epi-illumination microscope. The
  red arrows indicate the four illumination polarization orientations---one for
  each intensity measurement. b) Uncertainty ellipses for the microscope in
  a). The ellipses indicate the relative size of the uncertainty in different
  directions, not the absolute size of the uncertainty. c) Solid-angle
  uncertainty for the microscope in a). The solid-angle uncertainty is a measure
  of the absolute size of the uncertainty in all directions. d) Median of the
  solid-angle uncertainty as a function of illumination and detection NA. e) MAD
  of the solid-angle uncertainty as a function of illumination and detection
  NA. The microscope configuration in a), b), and c) is indicated by a cross in
  d) and e).}
\label{fig:single-arm}
\end{figure}

Figures \ref{fig:symmetric-widefield} and \ref{fig:double-arm} show our results
for dual-view symmetric designs. We illuminate from one objective and
detect from the other for four polarization orientations, then we repeat the
polarization orientations with the same objectives in reversed roles. In Figure
\ref{fig:symmetric-widefield} we sweep through the NA of both objectives and the
angle between the objectives while considering steric constraints. Note how
adding a second view removes the high uncertainty region in the transverse
plane. The dual-view microscope still has high uncertainty regions near the
optical axes, but the uncertainty is reduced by almost two orders of magnitude
compared to the single-view microscope. We find that the lowest median of the
solid-angle uncertainty occurs with the largest possible NA objectives. We also
find that increasing NA always lowers the MAD, but orthogonal arms are not
always optimal. At large NA, it is advantageous to move the objectives together
or against the cover slip. Given the high uncertainty in the transverse plane
for a single-view shown in Figure \ref{fig:single-arm}c), we can see why oblique
views perform better than orthogonal views---oblique views are complementary
because the regions of high orientation uncertainty from each view do not
overlap.

\begin{figure}[htbp]
\centering\includegraphics[width=\textwidth]{symmetric-widefield}
\caption{Dual-view symmetric designs with varying NA and angle between
  objectives. a) Schematic of the microscope. We illuminate with the first
  objective (red solid) and detect from the second objective (red dashed). Then
  we illuminate from the second objective (blue solid) and detect from the first
  objective (blue dashed). b) Uncertainty ellipses for the microscope in a). The
  ellipses indicate the relative size of the uncertainty in different
  directions, not the absolute size of the uncertainty. c) Solid-angle
  uncertainty for the microscope in a). The solid-angle uncertainty is a measure
  of the absolute size of the uncertainty in all directions. d) Median of the
  solid-angle uncertainty as a function of NA and the angle between the
  objectives. e) MAD of the solid-angle uncertainty as a function of NA and the
  angle between the objectives. The microscope configuration in a), b), and c)
  is indicated by a cross in d) and e). f) Median of the solid-angle uncertainty
  as a function of the angle between the objectives when NA=0.6. g) MAD of the
  solid-angle uncertainty as a function of the angle between the objectives when
  NA=0.6. The profile in f) and g) is taken along the purple line in d) and e).}
\label{fig:symmetric-widefield}
\end{figure}

Figure \ref{fig:double-arm} shows our results when we used a dual-view symmetric
orthogonal design and varied the illumination and detection NA. Our results are
similar to the single-view case in Figure \ref{fig:single-arm}. We find that
optimal designs use a small illumination NA and a large detection NA. As we
discussed in Section \ref{excitation}, low-NA illumination is
approximately equivalent to light-sheet illumination for the task of estimating
the orientation of molecules. Therefore, the preference for low-NA
illumination also implies that dual-view light-sheet illumination geometries are
an excellent choice for uniformly reconstructing the orientation of single
molecules.

\begin{figure}[htbp]
\centering\includegraphics[width=\textwidth]{double-arm}
\caption{Dual-view symmetric orthogonal designs with varying
  illumination and detection NA. a) Schematic of the microscope. b) Uncertainty
  ellipses for the microscope in a). The ellipses indicate the relative size of
  the uncertainty in different directions, not the absolute size of the
  uncertainty. c) Solid-angle uncertainty for the microscope in a). The
  solid-angle uncertainty is a measure of the absolute size of the uncertainty
  in all directions. d) Median of the solid-angle uncertainty as a function of
  illumination and detection NA. e) MAD of the solid-angle uncertainty
  as a function of illumination and detection NA. The microscope configuration
  in a), b), and c) is indicated by a cross in d) and e).}
\label{fig:double-arm}
\end{figure}

Figure \ref{fig:asymmetric-double} shows our results for dual-view asymmetric
light-sheet illumination designs. We used light-sheet illumination on both sides
and kept the objectives orthogonal so that the light sheet from one objective
illuminates the focal plane of the other objective. Then, we swept through the
NA of one objective while maximizing the NA of the other objective so that both
objectives would touch the cover slip and each other
($\alpha_1 + \alpha_2 = 90^{\circ}$ in Figure \ref{fig:asymmetric-double}a)). We
swept through the NA asymmetry and the sample-exposure asymmetry while keeping
the total sample exposure constant, and found that symmetric designs are at a
local minimum of the median and MAD of the solid-angle uncertainty. We also
found that at extreme asymmetries where the exposure from the low-NA arm is much
larger than the exposure from the high-NA arm (top-left and bottom-right corners
of Figures \ref{fig:asymmetric-double}d) and e)) the median and MAD are
comparable to the symmetric light-sheet microscope. This means that if a very
high-NA objective is available and a low-NA objective can provide light sheet
illumination, this microscope will perform comparably with a dual-view
microscope with two medium-NA objectives. Note that this asymmetric exposure
configuration maximizes the use of the high-NA detection objective---if a
high-NA objective is available it should be used for detection as much as
possible. Figure \ref{fig:asymmetric-double} also shows that designs with
slightly asymmetric NA can trade off a low solid-angle uncertainty median for a
low solid-angle uncertainty MAD by changing the sample exposure ratio.

\begin{figure}[htbp]
  \centering\includegraphics[width=\textwidth]{asymmetric-double}
  \caption{Dual-view asymmetric light-sheet illumination designs with varying NA
    and sample exposure asymmetry. a) Schematic of the microscope. b)
    Uncertainty ellipses for the microscope in a). The ellipses indicate the
    relative size of the uncertainty in different directions, not the absolute
    size of the uncertainty. c) Solid-angle uncertainty for the microscope in
    a). The solid-angle uncertainty is a measure of the absolute size of the
    uncertainty in all directions. d) Median of the solid-angle uncertainty as a
    function of NA and sample exposure asymmetry. e) MAD of the solid-angle
    uncertainty as a function of NA and sample exposure asymmetry. The
    microscope configuration in a), b), and c) is indicated by a cross in d) and e).}
\label{fig:asymmetric-double}
\end{figure}

Table 1 shows a summary of our evaluation metrics for microscopes in each
class. The single-view microscope with a small illumination NA and high
detection NA has the lowest solid-angle uncertainty median but the largest max
and fourfold symmetry. The dual-view oblique symmetric with an
intermediate NA performs reasonably well on all three metrics, while the
dual-view orthogonal symmetric light-sheet microscopes with large NA perform
very well on all three metrics.

\begin{table}[ht!]
\centering
\caption{Comparison of designs in each class of microscope. All solid-angle
  uncertainty statistics are in steradians.}
\begin{tabular}{lllll}
  \toprule

  Microscope Type&$n$-fold&Max\{$\sigma_{\Omega}$\}&Median\{$\sigma_{\Omega}$\}&MAD\{$\sigma_{\Omega}$\}\\ 
    &Degeneracy&&&\\\midrule
Single-view (NA${}_\textrm{ill}$=0, NA${}_\textrm{det}$=1.1)&4&3.04$\times 10^{0}$&8.20$\times 10^{-4}$&1.48$\times 10^{-4}$ \\ 
  Dual-view oblique symmetric (NA=0.6, $\beta$=53${}^{\circ}$)&2&1.35$\times 10^{-1}$&4.49$\times 10^{-3}$&9.17$\times 10^{-4}$\\
Dual-view orthogonal symmetric light-sheet (NA=0.8)&2&6.23$\times 10^{-3}$&1.79$\times 10^{-3}$&1.65$\times 10^{-4}$\\  
Dual-view orthogonal symmetric light-sheet (NA=0.94)&2&3.13$\times 10^{-3}$&1.28$\times 10^{-3}$&1.14$\times 10^{-4}$\\
\bottomrule
\end{tabular}
\end{table}

\section{Discussion and Conclusions}\label{discussion}
In this work we have developed a model to predict the intensities measured by
polarized illumination microscopes in a wide range of geometries, developed
metrics to measure the ability of a microscope design to estimate the
orientation of single molecules, and used these metrics to compare a wide
range of microscope designs. Our main result is a short list of design
heuristics that can be used to design polarized-illumination single-molecule
orientation microscopes:
\begin{itemize}
\item Single-view microscopes benefit from a small illumination NA and a large
  detection NA with $\alpha \approx 65^{\circ}$.
\item Dual-view microscopes outperform single-view microscopes along several
  metrics. Dual-view microscopes have fewer degeneracies, lower median and MAD
  solid-angle uncertainty, and no high solid-angle uncertainty region in the
  transverse plane.
\item High-NA dual-view microscopes are optimal when the two views are
  oblique. Oblique views are complementary because their high uncertainty
  transverse planes do not overlap.
\item Symmetric NA dual-view orthogonal light-sheet microscopes provide a good
  compromise between cost, solid-angle uncertainty median, and solid-angle
  uncertainty MAD. Asymmetric NA dual-view orthogonal light-sheet microscopes
  can improve the solid-angle uncertainty median or MAD by changing the sample
  exposure ratio between the two views.
\end{itemize}

Our results are limited in several ways. First, we only consider single
molecules in homogeneous environments. We ignore light reflected from the cover
slip and any other inhomogeneities in the sample. If we considered light
reflected from the cover slip we expect that the results would put a lower
weighting on large detection NAs because we would collect more light. Second, we
only considered ideal, aplanatic, polarization-preserving objectives. In
practice, these conditions are not always satisfied and the power from high-NA
rays is apodized by a factor of $\frac{n}{\cos\Theta}$
\cite{nov2006}. Therefore, our results are most accurate for low-NA rays and get
progressively worse for high-NA rays, although our major conclusions would not
change if we included this apodization factor. The main advantage of ignoring
apodization is the tractability of the problem---we have provided approximate
closed-form solutions for the illumination and detection efficiencies that have
allowed us to draw useful design conclusions. Third, we have used a Poisson
noise model which ignores noise introduced by the detector. Detector noise can
be a large fraction of the total noise at low intensities, so including detector
noise would increase the informational value of high intensity
measurements. Therefore, we expect that including the effects of detector noise
would favor microscopes with a large detection NA. Finally, we have focused on
the task of estimating the orientation of molecules and ignored the equally
important task of estimating their location. The task of jointly estimating the
orientation and location of single molecules from multiple views is a topic for
future work.

To compare the microscopes in this paper we used a fixed sample exposure. We
think that this is the fairest way to compare microscopes with different numbers
of intensity measurements. However, in cases where sample exposure and time are
not an issue it may be more useful to compare microscopes with the same total
detection exposure per measurement. This comparison would favor microscopes with
many intensity measurements even more that the results in this paper because
equal detection exposure would improve photon counting statistics.

We have considered single- and dual-view microscopes in a variety of cases, but
more multiview microscopes can be analyzed with the framework we have
developed. Wu et. al. have developed a three-view design with a third objective
below the cover slip\cite{wu2016}. Light-field detection schemes are also a type
of multiview microscope\cite{levoy2006}, and we plan to extend our analysis to
this case in future work.

Although we have only considered the orientation determination of fixed single
molecules, polarized illumination techniques can be used to study the
orientation distribution of ensembles of dynamic molecules \cite{mehta2016,
  backer2016}. In future work we will investigate the ability of multiview
polarized illumination microscopes to estimate the orientation distribution of
ensembles of dynamic molecules.

We have only considered polarized illumination. In the introduction
we discussed why we prefer polarized illumination to polarized detection, but
there is information to be gained by adding polarized detection to the
techniques discussed in this paper. Adding a polarization splitter to the
detection arm can increase the information available for reconstructing the
orientation of a single molecule without increasing the acquisition time.

We conclude that multiview microscopes are useful tools for determining the
orientation of single molecules. Using simple design heuristics, we can design
polarized illumination multiview microscopes that can determine the orientation
of single molecules with a small and uniform orientation uncertainty.

\section*{Funding}
Support for this work was provided by the Intramural Research Programs of the
National Institute of Biomedical Imaging and Bioengineering.

\section*{Disclosures}
The authors declare that there are no conflicts of interest related to this article.

\end{document}